A significance test for forward stepwise model selection

### 2013-10-09 paper/paper.tex
Paper editing

\newcommand{\grad}{\nabla}
\newcommand{\V}{\mathcal{V}}\endlocaldefs
\newcommand{\K}{\mathcal{K}}
\newcommand\tf{\widetilde{f}}

@@

\todo Expand this section

\begin{itemize}
  \item Brief background, cite some original papers
  \item Stopping rules, cite new paper on lasso?
\end{itemize}

@@

So far we have left open the question of choosing among the models in
the forward stepwise sequence, i.e. when to stop stepping
forward. Some classical approaches can be posed as optimization
criteria and stop at the step minimizing
\begin{equation}
\begin{aligned}
\label{eq:subsetregress}
\frac{1}{2} \| y - X \beta_s \|_2^2 + \lambda P(\beta_s)
\end{aligned}
\end{equation}
where we have written $\{ \beta_s : s = 1, \ldots, max.steps \}$ as
the sequence of models outputted by forward stepwise. The function
$P(\beta)$ is a penalty on model complexity usually taken to be the
number of nonzero entries of $\beta$. Proposals for $\lambda$ include
2 ($C_p$, AIC \todo{cite}), $\log(n)$ (BIC of \cite{BIC}), and
$2\log(p)$ (RIC of \cite{RIC}). In the current paper we provide
accurate p-values when adding the first noise variable group in the forward
stepwise sequence, and it is natural to consider using these p-values
to choose a model. \cite{sequential:fdr} examined some stopping rules
using the asymptotic p-values of \cite{significance:lasso} and showed
their stopping rules control false discovery rate--the expected
proportion of noise variables among variables declared significant
\citep{fdr}. We explore this further in Section \ref{sec:simulations}.

@@

In a recent work \cite{significance:lasso} defined a \emph{covariance
  test statistic} for testing the significance of a variable entering
the model along the lasso solution path.  They derived a simple
asymptotic null distribution, proved a type of convergence under broad
``minimum growth'' conditions, and demonstrated in simulations that
the test statistic closely matches its asymptotic distribution even in
finite samples.  That work marked an important advance in the problem
of combining inference with model selection.  The current paper
extends some of their results to the group lasso (ref Yuan and Lin?).
In the process we had to derive an exact finite sample null
distribution for the test statistic (ref TLTT?).  We also show that
the techniques used to get these results can be used to do a forward
stepwise procedure related to the group lasso that adds groups of
variables (or factors) in each step.

\subsection{Background}
In the ordinary least squares setting, a significance test for a
single variable can be conducted by comparing the drop in residual
sums of squares (RSS) to a $\chi^2_1$ distribution. Similarly, when
adding a group of $k$ variables we can compare the drop in RSS to a
$\chi^2_k$ random variable. This generally does not work when the
variable to be added has been chosen by a method that uses the data,
and in particular it fails for forward stepwise procedures which add
the ``best'' (e.g. most highly correlated) predictor in each step. In
that case, the test statistic (drop in RSS) does not match the
theoretical null distribution even when the null hypothesis is
true. Lockhart et al. (ref LTTT) introduced a new test statistic based
on the knots in the lasso solution path. Writing $\hat
\beta(\lambda)$, the solution for a fixed value of $\lambda$, we need
the following facts (ref Tibs2012)

>>>>

\newcommand{\grad}{\nabla}
\newcommand{\V}{\mathcal{V}}\endlocaldefs
\newcommand{\K}{\mathcal{K}}
\newcommand\tf{\widetilde{f}}
\newcommand{\pen}{\mathcal{P}}

@@

@@

So far we have left open the question of choosing among the models in
the forward stepwise sequence, i.e. when to stop stepping
forward. Some classical approaches can be posed as optimization
criteria and stop at the step minimizing
\begin{equation}
\begin{aligned}
\label{eq:subsetregress}
\frac{1}{2} \| y - X \beta_s \|_2^2 + \lambda \pen(\beta_s)
\end{aligned}
\end{equation}
where we have written $\{ \beta_s : s = 1, \ldots, max.steps \}$ as
the sequence of models outputted by forward stepwise. The function
$\pen(\beta)$ is a penalty on model complexity usually taken to be the
number of nonzero entries of $\beta$. Proposals for $\lambda$ include
2 ($C_p$, AIC \todo{cite}), $\log(n)$ (BIC of \cite{BIC}), and
$2\log(p)$ (RIC of \cite{RIC}). In the current paper we provide
accurate p-values when adding the first noise variable group in the forward
stepwise sequence, and it is natural to consider using these p-values
to choose a model. \cite{sequential:fdr} examined some stopping rules
using the asymptotic p-values of \cite{significance:lasso} and showed
their stopping rules control false discovery rate--the expected
proportion of noise variables among variables declared significant
\citep{fdr}. We explore this further in Section \ref{sec:simulations}.

@@

In the ordinary least squares setting, a significance test for a
single variable can be conducted by comparing the drop in residual
sums of squares (RSS) to a $\chi^2_1$ distribution. Similarly, when
adding a group of $k$ variables we can compare the drop in RSS to a
$\chi^2_k$ random variable. This generally does not work when the
variable to be added has been chosen by a method that uses the data,
and in particular it fails for forward stepwise procedures adding
the ``best'' (e.g. most highly correlated) predictor in each step. In
that case, the test statistic (drop in RSS) does not match the
theoretical null distribution even when the null hypothesis is
true. \cite{significance:lasso} introduced a new test statistic based
on the knots in the lasso solution path. They derived a simple
asymptotic null distribution, proved a type of convergence under broad
``minimum growth'' conditions, and demonstrated in simulations that
the test statistic closely matches its asymptotic distribution even in
finite samples.  That work marked an important advance in the problem
of combining inference with model selection.  The current paper
extends some of their results to the group lasso \citep{grouplasso}.
In the process we had to derive an exact finite sample null
distribution for the test statistic, this appears in
\cite{tests:adaptive}.

Writing $\hat
\beta(\lambda)$, the solution for a fixed value of $\lambda$, we need
the following facts (ref Tibs2012)
